CFF GCT Study

### 2024-08-27 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
GCT table 1

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(pROC)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| warning: false
# Import from REDCap
df <- read_excel("./Christine Chan/CFF GCT Study/Data_Raw/CFF Study REDCap Export 2021-09-09a LSP.xlsx")
# Get diagnosis
df$Diagnosis <- "NGT"
df$Diagnosis[df$plas_pre_gluc >= 100] <- "IFG"
df$Diagnosis[df$plasma_gluc_1_hr >= 200] <- "INDET"
df$Diagnosis[df$plas_gluc_2_hr >= 140] <- "IGF"
df$Diagnosis[df$plas_pre_gluc >= 126 | df$plas_gluc_2_hr >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD")
)
df$Dysglycemia <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
)
# Numeric variables
df$plasma_glu_val <- as.numeric(df$plasma_glu_val)
df$plasma_glucose <- as.numeric(df$plasma_glucose)
# Data were double entered but it wasn't completed so most participants have 3
# rows
df <- df %>%
  arrange(subject_id...1) %>%
  filter(grepl("--1|--2|-1|-2", subject_id...1)) %>%
  mutate(id = sub("--1|--2|-1|-2", "", subject_id...1)) %>%
  group_by(id) %>%
  filter(row_number() == 1) %>%
  select(
    id, gender2:ethnic_origin_race___6, age_at_enrollment_years,
    pancreatic_exocrine_insuff, Diagnosis, Dysglycemia,
    plasma_glucose, capillary, plasma_glu_val, hour_capillary_glucose_val,
    hemo_a1c, plas_pre_gluc, plasma_gluc_1_hr, plas_gluc_2_hr
  )
```

>>>>

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(pROC)
library(gtsummary)
library(gt)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

```{r data cleaning}
#| warning: false
# Import from REDCap
df <- read_excel("./Christine Chan/CFF GCT Study/Data_Raw/CFF Study REDCap Export 2021-09-09a LSP.xlsx")
# Get diagnosis
df$Diagnosis <- "NGT"
df$Diagnosis[df$plas_pre_gluc >= 100] <- "IFG"
df$Diagnosis[df$plasma_gluc_1_hr >= 200] <- "INDET"
df$Diagnosis[df$plas_gluc_2_hr >= 140] <- "IGF"
df$Diagnosis[df$plas_pre_gluc >= 126 | df$plas_gluc_2_hr >= 200] <- "CFRD"
df$Diagnosis <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD")
)
df$Dysglycemia <- factor(df$Diagnosis,
  levels = c("NGT", "IFG", "INDET", "IGF", "CFRD"),
  labels = c("NGT", "AGT", "AGT", "AGT", "AGT")
)
# Numeric variables
df$plasma_glu_val <- as.numeric(df$plasma_glu_val)
df$plasma_glucose <- as.numeric(df$plasma_glucose)
# Data were double entered but it wasn't completed so most participants have 3
# rows
df <- df %>%
  arrange(subject_id...1) %>%
  filter(grepl("--1|--2|-1|-2", subject_id...1)) %>%
  mutate(id = sub("--1|--2|-1|-2", "", subject_id...1)) %>%
  group_by(id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    id, gender2:ethnic_origin_race___6, age_at_enrollment_years,
    pancreatic_exocrine_insuff, Diagnosis, Dysglycemia,
    plasma_glucose, capillary, plasma_glu_val, hour_capillary_glucose_val,
    hemo_a1c, plas_pre_gluc, plasma_gluc_1_hr, plas_gluc_2_hr
  )
```

# Demographics

```{r}
#| label: table1a
#| tbl-cap: Participant characteristics by diagnosis
#| message: false
df %>%
  select(
    age_at_enrollment_years, gender2:ethnic_origin_race___6,
    pancreatic_exocrine_insuff, Diagnosis
  ) %>%
  tbl_summary(by = Diagnosis) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

```{r}
#| label: table1b
#| tbl-cap: Participant characteristics by dysglycemia
#| message: false
df %>%
  select(
    age_at_enrollment_years, gender2:ethnic_origin_race___6,
    pancreatic_exocrine_insuff, Dysglycemia
  ) %>%
  tbl_summary(by = Dysglycemia) %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```
